---
title: "Data Preparation - Modelling"
output: html_document
---




#### Set up


```{r Set up}

#clean start:
rm(list = ls());gc()

#libraries
library(stringr)      #For some variable creation
library(lubridate)    #For working with dates
library(openxlsx)     #For converting Excel's numerics to date-times
library(readxl)       #For reading excel files
library(plyr)         #for rbind.fill
library(purrr)        #For map functions
library(dplyr)

#file paths
base_path <- "C:/Users/jtodd/OneDrive - Bond University/LBP/AJA/"
raw_data_path <- paste0(base_path, "Data/Extracted/")
proc_data_path <- paste0(base_path, "Data/Processed/")
analysis_data_path <- paste0(base_path, "Data/Pre-Analysis/")

```



#### Data Preparation

Data Preparation:
* Load data with tweaks
* Remove presentations of individuals under the age of 18
* Standardise dates and times



```{r Load FN Data}
#Load FirstNet data
fn_data <- read.csv(file = paste0(proc_data_path, "FirstNet_Presentations.csv"), 
                    stringsAsFactors = FALSE)
fn_meds <- read.csv(file = paste0(proc_data_path, "FirstNet_Medications.csv"),
                    stringsAsFactors = FALSE)
fn_physio <- read_excel(path = paste0(raw_data_path, "FirstNet Physio 2019-2021 deidentified.xlsx"))

#Physio data requires brief tweaking to match the processed files
fn_physio$combined_id <- paste0(fn_physio$`Project ID`, "-", fn_physio$`ED Arrival Date/Time`)
fn_physio <- fn_physio[fn_physio$`Project ID` != 76,]
```

```{r Filtering - Patients under 18}
fn_data <- subset(fn_data, Age >= 18)
```



```{r Recoding - Inpatient and ED Times}

#Fixing the INP.Admit.date.time and INP.Discharge.date.time for FirstNet
fn_data$INP.Admit.date.time <- convertToDateTime(as.numeric(fn_data$INP.Admit.date.time))
fn_data$INP.Discharge.date.time <- convertToDateTime(as.numeric(fn_data$INP.Discharge.date.time))

#Reformat to be lubridate classes
fn_data$INP.Admit.date.time <- ymd_hms(fn_data$INP.Admit.date.time)
fn_data$INP.Discharge.date.time <- ymd_hms(fn_data$INP.Discharge.date.time)

#Also ED times for FirstNet
fn_data$ED.Arrival.Date.Time <- ymd_hms(fn_data$ED.Arrival.Date.Time)
fn_data$ED.Departure.date.time <- ymd_hms(fn_data$ED.Departure.date.time)

#Also change FirstNet orders data: need order date to be a date
fn_meds$Ordered.Date <- ymd_hms(fn_meds$Ordered.Date)


#Adjust the times at which consultations were requested
fn_physio$`Consult requested date/time` <- convertToDateTime(as.numeric(fn_physio$`Consult requested date/time`))
fn_physio$`Consult requested date/time` <- ymd_hms(fn_physio$`Consult requested date/time`)
fn_physio$Consult_Req_Time <- fn_physio$`Consult requested date/time`

```


```{r Recoding - LOS}
#Modify the LOS fields that only occasionally have values - need to convert to numeric
fn_data$ED.SS.LOS.Hrs <- ifelse(fn_data$ED.SS.LOS.Hrs == "NULL", 0, as.numeric(fn_data$ED.SS.LOS.Hrs))
fn_data$INP.LOS.Hrs <- ifelse(fn_data$INP.LOS.Hrs == "NULL", 0, as.numeric(fn_data$INP.LOS.Hrs))
```


#### Feature Construction

* Define the outcome variable (inpatient admission)
* Define the candidate independent variables

```{r Dependent Variable - Inpatient Admission}
#If inpatient los is at least one hour longer than short stay, treating that as a real inpatient stay
fn_data$Inp_Outcome <- ifelse(fn_data$Departure.destination == "INP" & fn_data$INP.LOS.Hrs > (fn_data$ED.SS.LOS.Hrs + 1),
                              1, 0)

```

Variables:

Demographic:
- Age (continuous)
- Gender (binary)

Presentation timing
- OutsideWorkingHours (binary)
- Weekend (binary)

Mode of arrival
- QAS_Arrival (binary)

Type of back pain
- LBP_Type (ternary)
- LBP_Sinister (binary)
- LBP_Radicular (binary)
- LBP_NS (binary)

Imaging
- Imaging_Any (binary)
- Imaging_CT (binary)
- Imaging_XR (binary)
- Imaging_MR (binary)

Pain Medication
- Med_Analgesic (binary)
- Med_NSAID (binary)
- Med_Opiate (binary)
- Med_Relaxant (binary)
- Med_Benzo (binary)
- Med_Antiepileptic (binary)
- Med_Antidepress (binary)
- Med_OralCortico (binary)

Other
- Physio_Review (binary)
- ED_los (continuous)
- ATS_1_2 (binary)
- ATS_3 (binary)
- ATS_4_5 (binary)



```{r Independent Variables - Demographics}
#Define names:
demo_vars <- c("Age_cont", "Gender")

#Age
fn_data$Age_cont <- fn_data$Age

#Gender
fn_data$Gender <- fn_data$Gender #redundant, but explicit anyway
```


```{r Independent Variables - Timing}
#Define names:
timing_vars <- c("Weekend", "Not_working_hrs")

#Weekend presentation:
fn_data$Weekend <- 1*(wday(fn_data$ED.Arrival.Date.Time, label = T, abbr = F) %in% c("Saturday","Sunday"))

#Outside working hour pres:
fn_data$Not_working_hrs <- 1*(hour(fn_data$ED.Arrival.Date.Time) %in% 8:16)
fn_data$Not_working_hrs[hour(fn_data$ED.Arrival.Date.Time) == 17 & minute(fn_data$ED.Arrival.Date.Time) == 0] <- 1
fn_data$Not_working_hrs <- fn_data$Not_working_hrs * (1 - fn_data$Weekend)
fn_data$Not_working_hrs <- 1-fn_data$Not_working_hrs
```


```{r Independent Variables - Mode of Arrival}
#define names:
arrival_vars <- c("QAS_arrival")

#QAS arrival
fn_data$QAS_arrival <- 1*str_detect(string = tolower(fn_data$Arrival.Mode), pattern = "ambulance")
```


```{r Independent Variables - Diagnosis}
#define names
type_vars <- c("LBP_type", "LBP_Sinister", "LBP_Radicular", "LBP_NS")

#LBP type - needs a bit of work, but only need the final column I create
fn_nslbp_codes <- c("278862001","91957002","161891005","424270008",
                    "134407002","278860009","209571002","161894002",
                    "26538006","20021007","135860001","279039007",
                    "300956001","282766005","239880009","209565008",
                    "279040009","34789001","61486003","203095000",
                    "209548004","249921008","262965006","262975009")
fn_radicular_codes <- c("247366003","307176005","307177001","24300005",
                        "86269002","202794004","46960006","202735001",
                        "128196005","23056005","18347007")
fn_sinister_codes <- c("309083007",
                       "19284003","192970008","207957008","308758008",
                       "426646004","212360005","11437003","284062002",
                       "281933002","202728009","2304001","125871005",
                       "125608002","207986006","230614002","202731005",
                       "207993005", "44237008", "48956000", "65491009", 
                       "129166009")

fn_data$LBP_Sinister <- 1*str_detect(string = fn_data$ED.Diagnosis, 
                                      pattern = paste0(fn_sinister_codes, collapse = "|")) 

fn_data$LBP_Radicular <- (1-fn_data$LBP_Sinister)*str_detect(string = fn_data$ED.Diagnosis, 
                                       pattern = paste0(fn_radicular_codes, collapse = "|"))
fn_data$LBP_NS <- (1-fn_data$LBP_Radicular - fn_data$LBP_Sinister)*str_detect(string = fn_data$ED.Diagnosis, 
                                   pattern = paste0(fn_nslbp_codes, collapse = "|"))
fn_data$LBP_type <- ifelse(fn_data$LBP_Sinister == 1, "Sinister", 
                           ifelse(fn_data$LBP_Radicular == 1, "Radicular", "NSLBP"))



```



```{r Independent Variables - Imaging}
#define names
image_vars <- c("Imaging_Any", "Imaging_CT", "Imaging_XR", "Imaging_MR")

#Load in the data specifying the exact orders relevant for each imaging type
image_cats <- read.csv(file = paste0(raw_data_path, "Imaging Categories.csv"), stringsAsFactors = FALSE)
colnames(image_cats)[1] <- "ORDER_CATALOG_MR"

#MR/CT/XR imaging values:
MR_orders <- image_cats$ORDER_CATALOG_MR
CT_orders <- image_cats$ORDER_CATALOG_CT
XR_orders <- image_cats$ORDER_CATALOG_XR

#Clean up empty fields:
MR_orders <- MR_orders[nchar(MR_orders)>0]
CT_orders <- CT_orders[nchar(CT_orders)>0]
XR_orders <- XR_orders[nchar(XR_orders)>0]


#going to use a loop, making it easier to follow. 
fn_data[,c("Imaging_Any","Imaging_CT", "Imaging_XR", "Imaging_MR")] <- 2
for(i in 1:nrow(fn_data)) {
  
  #Current project (not presentation) ID and ED time window
  cur_proj_id <- fn_data$Project.ID[i]
  start_ed_time <- fn_data$ED.Arrival.Date.Time[i]
  end_ed_time <- start_ed_time + seconds(60*60*fn_data$ED.Total.LOS.Hrs[i]) 
  
  
  #Matching data from medication table for same presentation and during ED time (including SSU)
  cur_orders <- subset(fn_meds, 
                       (Project.ID == cur_proj_id) & (Ordered.Date < end_ed_time) & (Ordered.Date >= start_ed_time))
  
  #Check for matches in the order catalog field
  fn_data$Imaging_CT[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(CT_orders))
  fn_data$Imaging_MR[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(MR_orders))
  fn_data$Imaging_XR[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(XR_orders))
  fn_data$Imaging_Any[i] <- max(fn_data$Imaging_CT[i], fn_data$Imaging_MR[i], fn_data$Imaging_XR[i])
  
  
}



```



```{r Independent Variables - Pain Medication}
#define names
med_vars <- c("Med_Analgesic", "Med_NSAIDs", "Med_Opiate", 
              "Med_Relaxant","Med_Benzo", "Med_Antiepileptic", 
              "Med_Antidepress", "Med_OralCortico")


#Define the medications in each category:
Analgesics <- c("amiTRIPTYLine","buprenorphine","buprenorphine-naloxone","codeine",
                "dexamethasone",  "Diazepam","diclofenac","diclofenac topical",
                "fentanyl","gabapentin","HYDROmorphone","ibuprofen","ketamine",
                "ketOROLAC","methADONe","midazolam","morphine","naproxen",
                "orphenadrine","oxycodone","oxycodone-naloxone","paracetamol",
                "paracetamol-codeine","prednisone","pregabalin","propOFol",
                "tAPENTadol","tRAMadol") #Everything
NSAIDs <- c("diclofenac", "diclofenac topical", "ibuprofen", "ketOROLAC", "naproxen")
Opiates <- c("buprenorphine", "buprenorphine-naloxone", "codeine", "fentanyl", 
                  "HYDROmorphone", "ketamine", "methADONe", "morphine", "oxycodone",
                  "oxycodone-naloxone", "paracetamol-codeine", "propOFol", "tAPENTadol",
                  "tRAMadol")
Relaxants <- c("orphenadrine", "Diazepam")
Benzos <- c("Diazepam", "midazolam")
Antiepileptics <- c("amiTRIPTYLine", "gabapentin", "pregabalin")
Antidepressants <- c("amiTRIPTYLine")
Oral_corticos <- c("dexamethasone", "prednisone")

#Initialise variables, loop through each row of data and fill them in
fn_data[,med_vars] <- 2
for(i in 1:nrow(fn_data)) {
  
  #Current project (not presentation) ID and ED time window
  cur_proj_id <- fn_data$Project.ID[i]
  start_ed_time <- fn_data$ED.Arrival.Date.Time[i]
  end_ed_time <- start_ed_time + seconds(60*60*fn_data$ED.Total.LOS.Hrs[i]) 
  
  
  #Matching data from medication table for same presentation and during ED time (including SSU)
  cur_orders <- subset(fn_meds, 
                       (Project.ID == cur_proj_id) & (Ordered.Date < end_ed_time) & (Ordered.Date >= start_ed_time))
  
  #Check if there are any matches in the order catalog
  fn_data$Med_Analgesic[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(Analgesics))
  fn_data$Med_NSAIDs[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(NSAIDs))
  fn_data$Med_Opiate[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(Opiates))
  fn_data$Med_Relaxant[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(Relaxants))
  fn_data$Med_Benzo[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(Benzos))
  fn_data$Med_Antiepileptic[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(Antiepileptics))
  fn_data$Med_Antidepress[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(Antidepressants))
  fn_data$Med_OralCortico[i] <- 1*any(tolower(cur_orders$ORDER_CATALOG) %in% tolower(Oral_corticos))
  
}



```


```{r Independent Variables - Other}
#define names
other_vars <- c("Physio_Review", "ED_los", "ATS_1_2", "ATS_3", "ATS_4_5")



#Physio review
fn_data$Physio_Review <- 2 #Initialising
for(i in 1:nrow(fn_data)) {
  
  
  #Current presentation ID and ED time window
  cur_proj_id <- fn_data$Project.ID[i]
  start_ed_time <- fn_data$ED.Arrival.Date.Time[i]
  end_ed_time <- start_ed_time + seconds(60*60*fn_data$ED.Total.LOS.Hrs[i]) 
  
  
  #Matching data from medication table for same presentation and during ED time (including SSU)
  cur_data <- subset(fn_physio,
                     (`Project ID` == cur_proj_id)  & (Consult_Req_Time < end_ed_time) & (Consult_Req_Time >= start_ed_time))
  
  #Check for any physio consults and record
  fn_data$Physio_Review[i] <- 1*any(str_detect(string = tolower(cur_data$`AH Consult`), pattern = "physiotherapy"))
  
  
}

#ED LOS - total LOS hrs has good agreement with the approach of adding SS and Actual ED together, so just use total
fn_data$ED_los <- fn_data$ED.Total.LOS.Hrs


#ATS_1_2, ATS_4_5, ATS_3
fn_data$ATS_1_2 <- ifelse(fn_data$ATS <= 2, 1, 0)
fn_data$ATS_3 <- ifelse(fn_data$ATS == 3, 1, 0)
fn_data$ATS_4_5 <- ifelse(fn_data$ATS >= 4, 1, 0)


```



#### Save the prepared data

```{r}
#Reduce the data to the fields of interest
modelling_data <- fn_data[,c("Inp_Outcome", demo_vars, timing_vars, arrival_vars, 
                             type_vars,image_vars, med_vars, other_vars)]

#Save: 
saveRDS(object = modelling_data, file = paste0(analysis_data_path, "inpatient_model_data.rds"))


```


